---
title: "Covid91 vaccine study FinalFall2022"
author: "Danitza Nunez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
```{r}
malesdata=filter(FinalData,sex=="M")
```


$H_0$: The treatment has no effect on males. 

$H_A$: The treatment has a significant effective on males. 


### Graphical Description
```{r}
barchartGC(~infected + treatment,data=malesdata)
barchartGC(~infected + treatment,data=malesdata, type="percent")

```


The graphical description depicts that among the males that contracted COVID91, about twice as many received the placebo as received the COVID91 treatment. From this graph the treatment is effective for males. 



### Numerical Description 

```{r}
table1 <- xtabs(~infected + treatment, data=malesdata)
rowPerc(table1)
colPerc(table1)
```
The numerical data reflects that among the males that contracted COVID91, 68.44% received the placebo, and 31.56% of males received the the COVID91 Vaccine treatment. The data shows that the treatment is effective for males.  


### Inferential Results  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The Chi-squared test has a P-Value of 0.000000000005307 which is less than 0.05 meaning we reject the null hypothesis. which concludes that the vaccine does have an effect on males. 

The Fisher's Test has a p-value of 0.000000000002364 which is less than 0.05 meaning we reject the null hypothesis. We conclude that the treatment does have an effect on males. 


Males who received the placebo are 2.18 more likely to contract COVID91 than males the vaccine treatment.



### Discussion and Conclusion

Among the males who contracted COVID91 68.44% received the placebo, and 31.56% received the vaccine treatment. Based on the data collected among the males who received the placebo, the odds are that they are  2.18 times more likely to contract COVID91 than males who received the treatment.


## Females

```{r}
femalesdata=filter(FinalData,sex=="F")
```

$H_0$: The treatment has no effect on females.  
$H_A$: The treatment has a significant effect on females. 



### Graphical Description  
```{r}
barchartGC(~infected + treatment,data=femalesdata)
barchartGC(~infected + treatment,data=femalesdata, type="percent")
```


Based on the graph above, among the females who contracted Covid91 about 15% more of them received the Covid91 treatment as received the placebo. The graph suggest that the treatment is not effective for females.  



### Numerical Description
```{r}
table1 <- xtabs(~infected + treatment, data=femalesdata)
rowPerc(table1)
colPerc(table1)
```


According to the numerical data, among the females that contracted COVID91 42.82% received the placebo, and 57.18% of females received the treatment. The data demonstrates the treatment is not effective for females.  


### Inferential Results  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Chi Squared test has a p-value of 0.005747 which is less than 0.05 meaning we reject the null hypothesis. We conclude that the treatment does have an effect on females making them more susceptible to contracting Covid91. 
The Fisher's Test has a p-value of 0.004887 which is less than 0.05 meaning we reject the null hypothesis. We conclude that the treatment does have an effect on females.

Females who received the placebo are 0.74 as likely to contract Covid91 than females who received the treatment.  


### Discussion and Conclusion  

Based on the data above, among the females who contracted Covid91 42.82% received the placebo and 57.18% received the treatment. According to the data, among the females who received the placebo, the odds are that they are 0.74 times as likely to contract Covid91 than females who received the treatment, meaning yes the treatment has a negative effect therefore we reject the null hypothesis.  



## LGBTQ

```{r}
lgbtqdata=filter(FinalData,LGBTQ=="gay")
```

$H_0$: The treatment has no effect on LGBTQ community.  
$H_A$: The treatment has a significant effect on LGBTQ community.  


### Graphical Description  
```{r}
barchartGC(~infected + treatment,data=lgbtqdata)
barchartGC(~infected + treatment,data=lgbtqdata, type="percent")
```
  
Based on the graph above, among the LGBTQ community those who contracted Covid91 about more than 80% of them received the Covid91 treatment as received the placebo. The graph suggest that the treatment not effective for the LGBTQ community.  

```{r}






## Druggies


# Overall Results and Conclusions